
%=======================   Default Templete   ==================
\documentclass[a4paper]{article}


% file with some default definations
\input{structure.tex}
\usepackage{listings}
\lstset{language=Python, basicstyle=\normalsize\sffamily\linespread{0.8}, numbers=left, numberstyle=\small, stepnumber=1, numbersep=5pt}
\usepackage{fancyhdr}
\setlength{\parindent}{0pt}

\pagestyle{fancy}
\fancyhf{}
\lhead{\textbf{\NAME\ (\ANDREWID)}}
\chead{\textbf{Assignment \HWNUM}}
\rhead{\COURSE}


%==================Header details======================
\newcommand\NAME{Raghukul, Vibhor}
\newcommand\ANDREWID{160538, 160778}
\newcommand\HWNUM{4}
\newcommand\COURSE{CS648}
%======================================================

% available formatted sections:
% - COMMAND LINE ENVIRONMENT: \begin{commandline} \end{commandline}
% - FILE CONTENTS ENVIRONMENT: \begin{file}[optional filename, defaults to "File"]
% - NUMBERED QUESTIONS ENVIRONMENT: \begin{question}[optional title]
% - WARNING TEXT ENVIRONMENT(can also be used for note): \begin{warn}[optional title, defaults to "Warning:"]
% - INFORMATION ENVIRONMENT(can be used to mention given details): \begin{info}[optional title, defaults to "Info:"]

%===============================================================
\begin{document}

% start of Q1
\begin{question}
\textbf{Internalizing Min-cut Algorithm}
\end{question}
\subsection*{1.}

\subsection*{2.}
\textbf{Time Complexity :}\\
Each contract step takes $O(n)$ and we are contracting $\frac{n}{2}$ times so this results in $O(n^2)$ work.\\
Thus if the time complexity of the algorithm is $T(n)$ then,
$$T(n) = O(n^2) + 2T(\frac{n}{2})$$
$$T(n) = O(n^2) + O(\frac{n^2}{2}) + 2T(\frac{n}{4})$$
$$T(n) = O(n^2) + O(\frac{n^2}{2}) + O(\frac{n^2}{8}) + \dots$$
$$T(n) = O(n^2)$$
\textbf{Error Analysis :}\\
Let the probability that algorithm preserves the min-cut be $q(n)$.\\
As shown in class, probability that min-cut is preserved after $\frac{n}{2}$ contractions is at least $\frac{1}{4}$.
Thus, we can say that $q(n)$ is at least $\frac{1}{4}$ times the probability that at least one smaller recursive call of size $\frac{n}{2}$ preserves the min-cut.
So,
$$q(n) \geq \frac{1}{4}\left(1-\left(1-q\left(\frac{n}{2}\right)\right)^2\right)$$
$$q(n) \geq \frac{1}{4}\left(2 - q\left(\frac{n}{2}\right)\right)q\left(\frac{n}{2}\right)$$
$$q(n) \geq \frac{1}{2}q\left(\frac{n}{2}\right) - \frac{1}{4}q\left(\frac{n}{2}\right)^2$$
Now let $p(k) = q(2^k)$. So we have,
$$p(k) \geq \frac{1}{2}p\left(k-1\right) - \frac{1}{4}p\left(k-1\right)^2$$
Call the above inequality to be $I_1$.\\
Since $p(k) \leq 1\ \forall k$ we get $$p(k) \geq \frac{1}{2}p\left(k-1\right) - \frac{1}{4}p\left(k-1\right)$$
or,
$$p(k) \geq \frac{1}{4}p\left(k-1\right)$$
By the above inequality and $I_1$ we have,
$$p(k) \geq \frac{1}{2}p\left(k-1\right) - p(k)p\left(k-1\right)$$
Thus,
$$\frac{1}{p(k)} \leq \frac{2}{p(k-1)} + 2$$
Base case - $p(1) = 1$
$$\frac{1}{p(k)} \leq 2\left(\frac{2}{p(k-2)} + 2\right) + 2$$
$$\frac{1}{p(k)} \leq \frac{2^2}{p(k-2)} + 2 + 2^2$$
In general,
$$\frac{1}{p(k)} \leq \frac{2^i}{p(k-i)} + 2 + 2^2 + \dots + 2^i$$
So,
$$\frac{1}{p(k)} \leq \frac{2^{k-1}}{p(1)} + 2 + 2^2 + \dots + 2^{k-1}$$
$$\frac{1}{p(k)} \leq 2^{k-1} + 2(2^{k-1}-1)$$
$$\frac{1}{p(k)} \leq 2^k + 2^{k-1} - 2$$
Thus,
$$\frac{1}{q(n)} \leq n + \frac{n}{2} - 2$$
$$q(n) \geq \frac{2}{3n - 4}$$
So the probability that the algorithm preserves min-cut is at least $\frac{2}{3n - 4}$\\
To get error probability less than $\frac{1}{n^c}$ repeat our algorithm $cn\log(n)$ times.\\
Hence, the time complexity of the complete algorithm to get an inverse polynomial error bound is $O(n^3 \log(n))$.
\pagebreak
% start of Q2
\begin{question}[]
\textbf{A surprising problem from computational geometry}
\end{question}
\subsubsection*{1.}
We will construct the line segments in order they are to be selected in the permutation. Let's say the first line segment is a horizontal line segment of length 1 having its' left end at origin. Construct the second line segment such that it will intersect the extended first line segment on the right side, and makes a small positive angle with the horizontal. 
Construct the third line segment which will intersect both the previous extended line segments, this line segment makes just greater angle with the horizontal than the previous line segment.
Now consider that we have constructed $i-1$ line segments in this manner, now construct the $i^{th}$ line segment intersecting all previous line segments and making an angle just greater than the ${(i-1)}^{th}$ line segment with the horizontal. Do this till we get $n$ line segments.\\
\textbf{Note - }I am considering angle with the horizontal in the anticlockwise sense.\\
This procedure will result in $n\choose2$ points of intersection as every pair of line segments will have a distinct point of intersection if constructed this way.\\
Thus, there exist a set of $n$ segments and a permutation for which the above algorithm will cause $n\choose2$ points of intersection.
\subsubsection*{2.}

\pagebreak

% start of Q3
\begin{question}[]
\textbf{Internalizing Backward Analysis and Randomized Incremental Construction (RIC)}
\end{question}
A point $x_i, y_i$ is said to dominate another point $x_j, y_j$ if $x_i > x_j$ and $y_i > y_j$. \\
\textbf{Non dominant point: } A point is said to be non dominant if there is no point in $P$ which dominates it. \\
So we need to find an expected $O(n\log n)$ algorithm to compute all non-dominanted point in $P$. We'll solve this using \textit{Randomized Incremental Construction}, using conflict graphs. \\ \\

Some notations used in analysis are $G_i: $ denotes the conflict graph after $i^{th}$ step, $p_i: $ $i^{th}$ point in our incremental construction, $I_i: $ $i^{th}$ interval on $X$ axis.

\subsubsection*{Conflict Graph Details}
Our conflict graph to solve this problem is based on interval division. Formally at $i^{th}$ stage, conflict graph $C_i$ will store 2-way mapping between set of points and set of intervals (division of $X$ axis). \\ \\ 
If $I_1 = 0\rightarrow x_1, I_2 = x_1 \rightarrow x_2, I_k \ldots x_{k-1} \rightarrow x_{k}$ denote the intervals at $i^{th}$ stage ($x_k$ being the point in $P$ having largest $x$ coordinate), then our conflict graph will store which interval, point $p$ belongs to $\forall p \in P$. Also we store a list of all the points belonging to a particular interval. Lastly we'll also store the points which define this interval, ie $x_{i-1}, x_i\in P$ for $I_i, \forall i \in [1,k]$, and the interval height (defined by $H_{i}$ for $I_i$ interval, $H_i = \max y_j \text{ for } j \in [1,i]$). Note that these interval heights may change when we proceed incrementally.\\ \\
Some operations and their complexity analysis are:
\begin{itemize}
	\item \textbf{Locating interval of $p_i \in P: $} Well store explicit mapping of interval for each point, so this opearation will take $O(1)$ time.
	\item \textbf{Determining whether $G_{i+1}$ will be different from $G_i: $} Note that $G_{i+1}$ will change only when we change the intervals in our graph. Interval will change if y-coordinate of $p_{i+1}$  is greater than $H_{t}$, where $H_{t}$ is the height of interval at $i^{th}$ stage to which $p_{i+1}$ belong. We can get the interval to which $p_{i+1}$ belongs in $O(1)$, also its height in $O(1)$, we determining whether $G_{i+1}$ will be different will take $O(1)$ time.
	\item \textbf{Updating $G_{i+1}$: } We will update $G_{i+1}$ only if $p_{i+1}$ has height greater than interval height to which $p_{i+1}$ belong to (let this interval be $I_t$). In that well remove $I_t = (x_{t-1} \rightarrow x_t)$ and add 2 new intervals: $I^1_t, I^2_t$, such that $I^1_t  = (x_{t-1} \rightarrow x_{p_i}), I^2_t = (x_{p_i} \rightarrow x_t)$ and height of $I^1_t$ is equal to y-coordinate of $p_{i+1}$ and height of $I^2_t$ is equal to height of $I_t$. Note that we have to change the mapping for all the points which belonged to $I_k$, by convention $p_{i+1}\in I^1_t$. Also we need to update neighbors of $I^1_t, I^2_t$ using neighbor data of $I_t$. So in worst case there can be $n$ points whose mapping needs to be changed, ie worst case time = $O(n)$. 
\end{itemize}
One last thing before we proceed further is that we'll also store the point $p$ which is on the top right corner of a interval. We have seen above that this point $p$ caused creation of this interval.

\pagebreak

So using these details of implementation as background our algorithm to find non dominant points is:

\begin{algorithm}[H]
\SetAlgoLined
\KwIn{$P = p_1, p_2,\ldots p_n:$ set of n points in $2-D$ plane}
\KwOut{Set of non dominant points in $P$}
  $G_1 = \{ I = \{ I_1 \}, M \}$ (initial conflict graph, here $I_1 = (0\rightarrow x_{mx})$, where $x_{mx}$ is maximum $X$ coordinate of all points in $P$) \\
  /* \\
  In subsequent step $M$ will denote the map from set of points to set of intervals \\
  $I$ will denote the set of intervals \\
  $I_k$ will denote $k^{th}$ interval, this will store data for height and interval limit(on $X$ axis).\\
  $H[I_k]$ will denote height of $k^{th}$ interval.\\
  $I_k$ is defined by $x_{k-1} \rightarrow x_k$\\
  $D[I_k]$ will denote the generating point for interval $I_k$ \\
  */ \\
  \For{$i \leftarrow 1$ \KwTo $n$} {
  	$M[p_i] = I_1$\\
  }
  $I_1 = (0\rightarrow x_{mx})$
  $H[I_1] = 0$\\
  $D[I_k] = p_{x_{mx}}$ // defined by point having maximum value of $X$ coordinate.
 \For{$i \leftarrow 1$ \KwTo $n$} {
 	$I_k$ = $M[p_i]$ \\
 	\uIf{$y_{p_i} > H[I_k]$}{
 			// have to update $G$ \\
 			$I^1_k = x_{k-1} \rightarrow x_{p_i}$ \\
 			$I^2_k = x_{p_i} \rightarrow x_k$ (since $I_k = x_{k-1} \rightarrow x_k$) \\
  			 \For{$j \leftarrow 1$ \KwTo $n$} {
  			 	\uIf{$M[p_j] = I_k$}{
  			 		\uIf{$x_{p_j} \geq x_{k-1} \& x_{p_j} \leq x_{p_i} $} {
  			 			$M[p_j] = I^1_k$ \\
  			 		}
  			 		\uElse {
  			 			$M[p_j] = I^2_k$ \\
  			 		}
  			 	}
 			 }
 			 $H[I^1_k] = y_{p_i}$ \\
 			 $H[I^2_k] = H[I_k]$ \\
 		}
 }
 $S = \phi$ \\
 \For{$Q \in I$} {
 	$S = S \cup \{ D[Q] \}$ // include defining point of this interval
 }

 \Return{$S$}
 \caption{RIC based algorithm for finding Non dominant points.}
\end{algorithm}


\end{document}

